### 2016-07-25 Reply_Referee1.tex
Alter second to last paragraph of the introduction to start to address comment #2  from referee 1 in the main paper. Adds responses to comments 2, 3 and 4 from Referee1 to the reply file, but the response to comment #2 needs review.

\documentclass[12pt,fleqn]{article}

% Add commit information from Git to the pdf. It is automatically
% generated with the R script 'tex/git and can be run from
% the command line in the project's top directory via:
%
% $ tex/git
%
% If VERSION.tex does not exist we can't add information from
% Git, so we'll use today's date as a fallback.

\IfFileExists{./VERSION.tex}{\input{VERSION}}{%
\providecommand\VERSION{\today}}

\input{tex/setup}
\input{tex/macros}

\title{Reply to the report by Referee 1 for ``Bootstrap Confidence Intervals for Sharp Regression Discontinuity Designs
  with the Uniform Kernel''}

\author{Ot\'avio Bartalotti \and Gray Calhoun \and Yang He\thanks{%
  All authors: Department of Economics, Iowa State University.
  260 Heady Hall, Ames, IA 50011.
  Bartalotti: \protect\url{bartalot@iastate.edu};
  Calhoun: \protect\url{gcalhoun@iastate.edu} and
  \protect\url{http://gray.clhn.org};
  He: \protect\url{yanghe@iastate.edu}.}}

\begin{document}
\maketitle

First, we thank the reviewer for providing detailed and helpful comments, which we feel have significantly improved the quality of the draft.  Below we provide replies to the reviewer’s specific comments detailing how we address each comment.

\section{Main Comments}

\begin{enumerate}
 \item \textit{``...the proposed bootstrap method is only valid under conditional homocedasticity for distributional approximations...''}

  We thank the referee for pointing out this limitation of the procedure proposed initially and suggesting the use of a wild bootstrap. In the new version of the paper we provide a wild bootstrap algorithm and results about its properties, allowing for conditional heteroskedasticity.

\item \textit{``...the need of a computationally intensive iterated bootstrap for distributional approximations...''}

   We share the reviewer's concern about the computing power demands an iterative bootstrap procedure imposes. Even though it would certainly be desirable to obtain an alternative implementation method that could achieve similar outcomes with a lesser burden on computational resources, it is not immediately obvious to the authors what that approach would be, and that was not the focus of the advances being proposed currently in this paper. Additionally, the simulation and computational work developed in this study seem to indicate that, in practice, the iterated bootstrap procedure proposed is not excessivelly burdensome for most applications faced by practitioners using standard computing resources widely available. Clearly, while we embrace the merit and usefulness of developing less intensive methods, we believe that even a demanding procedure like ours can be easily implemented by practitioners.
\end{enumerate}

\section{Specific Comments}

\begin{enumerate}
 \item  \textit{``It is not clear what is special about the uniform kernel. If possible, the authors should consider a general kernel and remove Uniform kernel from the title.''}

 We added results that allow for more general kernels typically used in RDD. The title has been changed accordingly.

\item \textit{``...What are the potential benefits of bootstrap over the analytical calculations.''}

   \textbf{We need to decide what to answer here. In the current version of the paper (which I assume is what they read) we have a paragraph discussing it that needs to be updated with the kernel and wild bootstrap innovations. However it is not clear to me what else he wants or what we could write. Here is the full original paragraph. I will start editing it in the main file, but feel free to add your two cents if you have an idea. ``The methods presented in this paper are intended as a ``proof of concept'' that
demonstrates that the analytical correction proposed by CCT can also be
implemented through a new bootstrap procedure. We believe that this procedure is
promising because it can allow applied researchers to construct
accurate confidence intervals in settings where unusual forms of dependence
would make it difficult to derive CCT's analytical correction---the dependence
can be accommodates in principle by modifying the resampling algorithm to one
that matches the dependence in the dataset. The specific version of the bootstrap
presented in this paper, however, does not achieve that level of generality yet,
and two limitations should be noted in particular. First, the confidence
intervals produced by this bootstrap depend on two separate bandwidth parameters
that must be generated elsewhere. Since these bandwidths are generated using
analytical methods, the confidence intervals could easily be produced using the
same methods. Second, the bootstrap proposed in this paper imposes local
homoskedasticity and can perform poorly when the DGP exhibits significant
heteroskedasticity. Moreover, in this paper we focus on the baseline case of
sharp RD with a local linear model and uniform kernel. So the practical benefits
of using the bootstrap in this context are relatively limited. Relaxing these
limitations is the subject of ongoing research by the authors.''}

\item \textit{`` The most critical point: Assumption 1.4 allows for heteroskedasticity but I suspect the residual bootstrap is invalid in this setting...''}

 We thank the reviewer for focusing our attention to this point and suggesting the use of the Wild bootstrap. Following his recommendation we altered the algorithm to use a wild bootstrap approach and discuss its properties, hence allowing for conditional heteroskedasticity.

\item \textit{``Are the bootstrap residuals conditionally centered?...''}

 Yes, the bootstrap residuals are conditionally centered. Hopefully this is clear on the discussion about the wild bootstrap approach.

\end{enumerate}

\end{document}