### 2016-08-24 reply_2/cover_letter.tex
[#22] Add discussion of sign error in simulations to cover letter
We should probably mention the fix to Matias. Done :).  Closes #22.

\documentclass[12pt,fleqn]{article}

% Add commit information from Git to the pdf. It is automatically
% generated with the R script 'tex/git and can be run from
% the command line in the project's top directory via:
%
% $ tex/git
%
% If VERSION.tex does not exist we can't add information from
% Git, so we'll use today's date as a fallback.

\IfFileExists{../VERSION.tex}{\input{../VERSION}}{%
\providecommand\VERSION{\today}}

\input{../tex/setup}
\input{../tex/macros}

\begin{document}

\noindent%
Hi Matias,

\strut

In addition, we discovered a small error in the simulation results ---
instead of reporting each estimator's bias, we reported the negative
bias. Since the values were close to zero, this does not have a
substantive effect on the results but we have, of course, fixed the
tables.

\noindent%
Best,

\noindent%
Ot\'avio Bartalotti, Gray Calhoun, and Yang He

\end{document}